---
title: "Lab 01: Hello R!"
subtitle: "Due Friday July 1 at 11:59pm"
author: ""
date: ""
output:
  html_document: default
  pdf_document: default
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(datasauRus)
``` 

# Exercise 1

Type your answer to exercise 1 here. Note this exercise does not require any
R code. 

Note also in this lab we are telling you exactly when to stage, commit, and
push.

# Exercise 2

The answer for this exercise is given below, but you should clean up the 
narrative (that's this text here!) so it only includes what you want to turn in.

First, let's plot the data in the dino data set  (we will cover `ggplot2` in
much more depth later):

```{r plot-dino}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")
ggplot(data = dino_data, mapping = aes(x = x, y = y)) + 
  geom_point()
```

Next calculate the correlation between `x` and `y` in this data set 
(an explanation for this code is given in the lab document).

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

# Exercise 3

Add code and narrative here as needed. Note the code chunks below are labeled
with `ex-3` and should be changed.

When you have finished, delete these instructions.

```{r ex3-1}
# Your code should go here! This is a comment, which should be deleted.
```

Some more narrative can go here.

```{r ex3-2}
# Calculate the correlation here
```

Conclude with some more narrative, if needed.

# Exercise 4

Add a labeled code chunk and narrative here. Insert code chunks using the 
"insert chunk" button (a green C with a +) and select `R`. Alternatively, 
use CMD + OPTION + I (Mac) or CTRL + ALT + I (Windows/Linux).

Be sure to name your code chunks!

# Exercise 5

You can run the code in the console, or create a new code chunk. Delete this narrative when you are done.

# Exercise 6

Use the space provided to label each. Delete this narrative when you are done.

- `summarize`: 

- `dino_data`: 

- `mean`: 

- `x`: 

- `y`: 

- `mu_x = mean(x)`: 

# Exercise 7

Add a labeled code chunk and narrative here. Insert code chunks using the 
"insert chunk" button (a green C with a +) and select `R`. Alternatively, 
use CMD + OPTION + I (Mac) or CTRL + ALT + I (Windows/Linux).

Be sure to name your code chunks!


