---
title: "Welcome to R"
subtitle: "STA 101"
format: html
editor: visual
---

## Today

By the end of today you will...

-   begin to know your way around RStudio
-   be able to define package, data frame, variable, function, argument
-   use the function `glimpse()`

## Getting started

Download this application exercise by pasting the code below into your **console** (bottom left of screen)

    download.file("https://sta101-fa22.netlify.app/static/appex/ae1.qmd",
    destfile = "ae1.qmd")

## R as a calculator

-   Use `R` as a calculator by typing the following into the console:

```{=html}
<!-- -->
```
    5 * 5 + 10

    x = 3
    x + x^2

    x = 1:10
    x * 7

In the last couple examples we save some value as the object "x".

We can "print" x to the screen by typing the name of the object ("x") in the console or in a code chunk.

## Tour of RStudio

-   environment
-   `R` functions
-   loading and viewing a data frame

## Load a package

```{r load-package, warning=F, message=F}
library(tidyverse) 
```

## Load data

```{r load-data, warning=F, message=F}
roster = read_csv("https://sta101.github.io/static/appex/data/sample-roster.csv")
survey = read_csv("https://sta101.github.io/static/appex/data/sample-survey.csv")
```

**Question**: What objects store the data in the code chunk above? Can you print them to the screen?

Create a new code chunk with `CMD+OPTION+I` (mac) or `CTRL+ALT+I` (windows/linux)

So far we've already seen two functions. `library` and `read_csv`. **Functions** in R are attached to parentheses and take an input, aka an **argument**, and often (but not always) return an output. To learn more about a function, you can check the documentation with `?`, e.g. `?library`.

## Demos

Let's glimpse the data frame.

```{r glimpse-data}
glimpse(survey)
```

To look at all of it, we can use `view()`

```{r view-data}
view(survey)
```

**View the roster data in the console**

Terminology: "columns" of a dataframe are called **variables** whereas "rows" are **observations**.

**Question: How many variables are in the data frame `survey`? How many observations? What about the data frame `roster`?**

Why must I input a specific email format?

```{r join-roster, eval=F}
roster %>% 
  left_join(survey, by = "email")
```
