---
title: "Welcome to R"
subtitle: "STA 101"
output: html_document
date: "June 27, 2022"
---

## Today

By the end of today you will... 

- begin to know your way around RStudio
- be able to define package, data frame, variable, function, argument
- use a few functions, including: `glimpse()`, `count()`, `group_by()` and `summarize()`
- play with data

## Getting started

Download this application exercise by pasting the code below into your **console** (bottom left of screen)

```{r download-rmd, eval=F}
download.file("https://sta101.github.io/static/appex/ae1.Rmd",
destfile = "ae1.rmd")
```

## R as a calculator

- Use `R` as a calculator by typing the following into the console:

```
5 * 5 + 10
```

```
x = 3
x + x^2
```

```
x = 1:10
x * 7
```

In the last couple examples we save some value as the object "x".

We can "print" x to the screen by typing the name of the object ("x") in the console or in a code chunk.

## Tour of RStudio

- environment
- `R` functions
- loading and viewing a data frame
- accessing a variable in a data frame


## Load a package

```{r load-package, warning=F, message=F}
library(tidyverse) 
```

## Load data
```{r load-data, warning=F, message=F}
roster = read_csv("https://sta101.github.io/static/appex/data/sample-roster.csv")
survey = read_csv("https://sta101.github.io/static/appex/data/sample-survey.csv")
```

**Question**: What objects store the data in the code chunk above? Can you print them to the screen?

Create a new code chunk with `CMD+OPTION+I` (mac) or `CTRL+ALT+I` (windows/linux)

So far we've already seen two functions. `library` and `read_csv`. **Functions** in R are attached to parentheses and take an input, aka an **argument**, and often (but not always) return an output. To learn more about a function, you can check the documentation with `?`, e.g. `?library`.



## Demos

Let's glimpse the data frame. 

```{r glimpse-data}
glimpse(survey)
```

To look at all of it, we can use `view()`

```{r view-data}
view(survey)
```

**View the roster data in the console**

Terminology: "columns" of a dataframe are called **variables** whereas "rows" are **observations**.

**Question: How many variables are in the data frame `survey`? How many observations? What about the data frame `roster`?**

Why must I input net-id email?

```{r join-roster, eval=F}
roster %>% 
  left_join(survey, by = "email")
```

Count the benders in the data

```{r count-benders}
count(survey, bender)
```


```{r count-cat-dog-by-group}
survey %>%
  mutate(pet = ifelse(cat_dog == "dog", 1, 0)) %>%
  group_by(bender) %>%
  summarize(proportion_dog = mean(pet))
```

