---
title: "Central limit theorem"
output: html_document
date: "July 19, 2022"
---


## Bulletin

- Regression project due Wednesday (tomorrow) 11:59p
- No lab today
- Exam 2 Thursday
- Review session tomorrow 4-5p in this room

## Today

By the end of today you will...


- Use Central Limit Theorem to define distribution of sample means
- Calculate probabilities from the normal distribution
- Use Central Limit Theorem (CLT) to conduct inference on a population mean

## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101.github.io/static/appex/ae10.Rmd",
destfile = "ae10.rmd")
```

## Load packages

```{r load-libraries, message = F, warning = F}
library(tidyverse)
library(tidymodels)
```


```{r load-data, message = F, warning = F}
manhattan = read_csv("https://sta101.github.io/static/appex/data/manhattan.csv")
```


## Notes

How do we know when to expect a normal distribution to show up?

Last time we saw an example where the distribution of **sample means** was normal while the distribution of **sample medians** was not.

Example:

```{r}
set.seed(1)
boot_dist = manhattan %>%
  specify(response = rent) %>% 
  generate(reps = 1000, type = "bootstrap")

boot_dist %>%
  calculate(stat = "mean") %>%
  visualize() +
  labs(x = "Sample mean", 
       title = "Simulated distribution of the sample mean")

boot_dist %>%
  calculate(stat = "median") %>%
  visualize() + 
  labs(x = "Sample median",
       title = "Simulated distribution of the sample median")
```

Are there times when the sample mean will not look normal?

Returning to our binomial example from `ae8`: you ask `numPeople` if they prefer dogs or cats. You repeat this 1000 times and plot the distribution of the proportion of who prefer dogs.

#### Exercise 1 

How does the shape of the distribution change as you increase the number of people per sample?

```{r}
set.seed(714)
numPeople = 1
numDog = rbinom(n = 1000, size = numPeople, prob = 0.85)
sample = data.frame(numDog) # new data frame called sample
sample %>%
mutate(propDog = numDog / numPeople) %>%
ggplot(aes(x = propDog)) +
geom_histogram(binwidth = .01)
```

## What is the central limit theorem?

The central limit theorem is a statement about the distribution of the sample mean, $\bar{x}$

The central limit theorem guarantees that, when certain criteria are satisfied, the sample mean ($\bar{x}$) is normally distributed.

Specifically,

If 

1. Observations in the sample are independent. Two rules of thumb to check this:

	- completely random sampling
	- if sampling without replacement, sample should be less than 10% of the population size

and

2. The sample is large enough. The required size varies in different contexts, but some good rules of thumb are: 

	- if the population itself is normal, sample size does not matter.
	- if numerical require, >30 observations
	- if binary outcome, at least 10 successes and 10 failures.

then 

$\bar{x} \sim N(\mu, \sigma / \sqrt{n})$

i.e. $\bar{x}$ is normally distributed (unimodal and symmetric with bell shape) with mean $\mu$ and standard deviation $\sigma / \sqrt{n}$. The standard deviation of the sampling distribution is called the **standard error**.

Note the standard deviation depends on the number of samples, $n$.


## Practice using CLT & Normal distribution

Suppose the bone density for 65-year-old women is normally 
distributed with mean $809 mg/cm^3$ and standard deviation of $140 mg/cm^3$.

Let $x$ be the bone density of 65-year-old women. We can write this distribution of $x$ in mathematical notation as 

$$x \sim N(809, 140)$$

## Visualize the population distribution

```{r visualize-normal}
ggplot(data = data.frame(x = c(809 - 140*3, 809 + 140*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140),
                color = "black") +
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140/sqrt(10)),
                color = "red",lty = 2) + theme_bw() +
  labs(title = "Black solid line = population dist., Red dotted line = sampling dist.")
```



#### Exercise 2

Before typing any code, based on what you know about the normal distribution, what do you expect the median bone density to be?

What bone densities correspond to $Q_1$ (25th percentile), $Q_2$ (50th percentile), and $Q_3$ (the 75th percentile) of this distribution? Use the `qnorm()` function to calculate these values. 

#### Exercise 3

The densities of three woods are below:

- Plywood: 540 mg/cubic centimeter
- Pine: 600 mg/cubic centimeter
- Mahogany: 710 mg/cubic centimeter

- What is the probability that a randomly selected 65-year-old woman has bones less dense than Pine?

- Would you be surprised if a randomly selected 65-year-old  woman had bone density less than Mahogany? What if she had bone density less than Plywood? Use the respective probabilities to support your response. 

#### Exercise 4

Suppose you want to analyze the mean bone density for a group of 10 randomly selected 65-year-old women. 

- Are the conditions for the Central Limit Theorem met?

  - Independence? 
  - Sample size/distribution? 
  
- What is the shape, center, and spread of the distribution of $\bar{x}$, the mean bone density for a group of 10 randomly selected 65-year-old women?

- Write the distribution of $\bar{x}$ using mathematical notation. 

#### Exercise 5

- What is the probability that the mean bone density for the group of 10 randomly-selected 65-year-old women is less dense than Pine?

- Would you be surprised if a group of 10 randomly-selected 65-year old women had a mean bone density less than Mahogany? What the group had a mean bone density less than Plywood? Use the respective probabilities to support your response. 

#### Exercise 6

Explain how your answers differ in Exercises 2 and 4.

## On your own

Suppose the distribution of the number of minutes users engage with apps on an iPad has a mean of 8.2 minutes and standard deviation of 1 minute. Let $x$ be the number of minutes users engage with apps on an iPad, $\mu$ be the population mean and $\sigma$ the population standard deviation. Then,

$$x \sim N(8.2, 1)$$

Suppose you take a sample of 60 randomly selected app users and calculate the mean number of minutes they engage with apps on an iPad, $\bar{x}$. The conditions (independence & sample size/distribution) to apply the Central Limit Theorem are met. Then by the Central Limit Theorem

$$\bar{x} \sim N(8.2, 1/\sqrt{60})$$

-   What is the probability a randomly selected user engages with iPad apps for more than 8.3 minutes? Use `pnorm` for calculations.

    ```{r review-1}
    #add code
    ```

-   What is the probability the mean minutes of app engagement for a group of 60 randomly selected iPad users is more than 8.3 minutes? Use `pnorm` for calculations.

    ```{r review-2}
    #add code
    ```

-   What is the probability the mean minutes of app engagement for a group of 60 randomly selected iPad users is between 8.3 and 8.4 minutes? Use `pnorm` for calculations. 

```{r review-3}
    #add code
```

## Data: Pokemon

We will be using the `pokemon` data set, which contains information about 42 randomly selected Pokemon (from all generations). You may load in the data set with the following code:

```{r load-data-2, warning = FALSE, message = FALSE}
pokemon = read_csv("https://sta101.github.io/static/appex/data/pokemon.csv")
```

In this analysis, we will use CLT-based inference to draw conclusions about the mean height among all Pokemon species.

#### Exercise 7

Let's start by looking at the distribution of `height_m`, the typical height in meters for a Pokemon species, using a visualization and summary statistics.

```{r height-histogram}
ggplot(data = pokemon, aes(x = height_m)) +
  geom_histogram(binwidth = 0.25, fill = "steelblue", color = "black") + 
  labs(x = "Height (in meters)", 
       y = "Distributon of Pokemon heights")
```

```{r height-stats}
pokemon %>%
  summarise(mean_height = mean(height_m), 
            sd_height = sd(height_m), 
            n_pokemon = n())
```

In the previous lecture (and in the review questions), we were given the mean, $\mu$, and standard deviation, $\sigma$, of the population. That is unrealistic in practice (if we knew $\mu$ and $\sigma$, we wouldn't need to do statistical inference!).

Today we will start on using the Central Limit Theorem to draw conclusions about the $\mu$, the mean height in the population of Pokemon.

-   What is the point estimate for $\mu$, i.e., the "best guess" for the mean height of all Pokemon?

-   What is the point estimate for $\sigma$, i.e., the "best guess" for the standard deviation of the distribution of Pokemon heights?

#### Exercise 8

Before moving forward, let's check the conditions required to apply the Central Limit Theorem. Are the following conditions met:

-   Independence?
-   Sample size/distribution?

## Idea: confidence intervals using CLT

```{r}

```

