---
title: "The Normal Distribution"
subtitle: "STA 101"
format: html
---

## Bulletin

- Regression project **due Friday October 14**

## Today

By the end of today you will...

- define **stable distribution**, **normal distribution**, **location** and **scale** parameters
- plot normal distributions and calculate probabilities
- 
## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae13.qmd",
destfile = "ae13.qmd")
```

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

## Normal distribution

The **normal distribution**, also known as "Gaussian distribution" is a distribution of a continuous random variable. The sample space of a normal random variable is $\{- \infty, + \infty \}$ and is defined by two parameters: a mean $\mu$ and a standard deviation $\sigma$. The mean is known as the **location** parameter while the standard deviation is the **scale**.

We can sample `N` times from a normal with mean `mu` and standard deviation `sigma` using `rnorm(n = N, mean = mu, sd = sigma)`.

```{r, eval = F}
# example
sample = rnorm(1000, mean = 0, sd = 1)

mean(sample)
sd(sample)

hist(sample)
```


Let's visualize the normal function curve using the code below.

```{r visualize-normal}

mu1 = 100
s1 = 6

mu2 = 105
s2 = 3

ggplot(data = data.frame(x = c(mu1 - s1*3, mu1 + s1*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = mu1, sd = s1),
                color = "steelblue") +
  stat_function(fun = dnorm, args = list(mean = mu2, sd = s2),
                color = "orange") + 
  theme_bw() +
  labs(title = "Two normal curves")
```

#### Exercise 1

Try setting `mu2` to 99 and `s2` to 4. What do you notice? Play with a few more settings and describe what the mean and standard deviation do to the shape of the curve.

#### Exercise 2

Start with the code below of a histogram from but change your geometry to `geom_histogram(aes(y = ..density..))`. This will re-scale your histogram so that the area under the curve is 1. Next, use the `stat_function` code above as a template to superimpose a normal distribution on top of your histogram. Adjust the mean and standard deviation until you obtain a good looking fit. What do you notice?

```{r eval=F}
set.seed(1013)
# todo: code here
```


## Properties of a Gaussian

If $X$ is a random variable, and $X$ is normally distributed, then it the distribution of $X$ is fully specified by the **location** (mean) and **scale** (standard deviation) parameters, $\mu$ and $\sigma$ respectively. In mathematical notation we write

$$
X \sim N(\mu, \sigma)
$$
and this reads: "X is normally distributed with mean mu and standard deviation sigma." It is worth noting that in **some contexts** (see e.g. "notation" on [wikipedia](https://en.wikipedia.org/wiki/Normal_distribution)) the second term of a Normal distribution refers to the variance or $\sigma^2$. 

A very useful property of Normal distribution is that it is **stable**. What this means is that linear combinations of normal random variables are themselves normal. In other words, if $X$ and $Y$ are normal random variables then $aX + bY$ is normal for all $a$ and $b$. The two properties to remember when adding random variables are:

- mean(X + Y) = mean(X) + mean(Y)
- variance(X + Y) = variance(X + Y) when X and Y are independent

We can see this in an example.

Let $X \sim N(5, 3)$ and let $Y \sim N(-5, 1)$

According to our rules above $X + Y \sim N(0, \sqrt{10})$. Let's check ourselves with code:

```{r}
set.seed(1)
normal_df = data.frame(X = rnorm(1000, mean = 5, sd = 3),
                 Y = rnorm(1000, mean = -5, sd = 1))



normal_df = normal_df %>%
  mutate(Z = X + Y)

normal_df %>%
  ggplot(aes(x = Z)) +
  geom_histogram(aes(y = ..density..), fill = 'gold3', alpha = 0.5) +
  theme_bw() +
  labs(y = "Density", title = "Matching densities") +
  stat_function(fun = dnorm, args = list(mean = 0, sd = sqrt(10)),
                color = "darkblue")
```

#### Exercise 3

Let $Z \sim N(0, 1)$ and $X \sim N(10, 4)$

$aZ + b \sim X$ for some $a$ and $b$. What are $a$ and $b$? Fill them in for the `?` in the code below and the uncomment the code.

```{r, eval = F}
set.seed(1)
normal_df = data.frame(z = rnorm(1000, mean = 0, sd = 1),
                 x = rnorm(1000, mean = 10, sd = 4))

sample_mean = normal_df %>%
  summarize(sample_mean = mean(x)) %>%
  pull()

normal_df %>%
  ggplot(aes(x = x)) +
  geom_histogram(aes(y = ..density..), fill = 'gold3', alpha = 0.5) +
  geom_histogram(aes(x = z, y = ..density..), fill = 'steelblue', alpha = 0.7) +
  #geom_density(aes(x = z*? + ?), color = 'red') + 
  theme_bw() +
  labs(x = "", y = "", title = "Sampling from two normals")
```

## Computing probabilities

`pnorm` "probability normal" takes three arguments:

- `q`, `mean` and `sd`

and answers the question:

If $X \sim N(\mu, \sigma)$, what is $p(X < q)$?

For example, imagine that the resting heart rates in the classroom are normally distributed with mean 70 beats per minute (bpm) and standard deviation 5 bpm. What's the probably a randomly selected individual has a heart rate less than 63 bpm?

In math: let $X$ be the bpm of an individual in this class $X \sim N(70, 5)$. What is $p(X < 63)$? Given heartbeats are normally distributed, randomly selecting an individual from the classroom is called "drawing from a normal distribution".

We can compute this easily:

```{r}
pnorm(63, 70, 5)
```

0.08 or about 8% chance. In picture, the probability is the proportion of area under the curve shaded:

```{r echo=F, message = F}
x=seq(50,90,length=200)
y=dnorm(x,mean=70,sd=5)
plot(x,y,type="l")
x=seq(30,63,length=100)
y=dnorm(x,mean=70,sd=5)
polygon(c(30,x,63),c(0,y,0),col="steelblue")

# mu1 = 70
# s1 = 5
# 
# shadeFrom = 50
# shadeLeft = shadeFrom - 0.0125
# 
# dnorm_limit <- function(x) {
#   y <- dnorm(x)
#   y[x < shadeLeft  |  x > 63] <- NA
#   return(y)
# }
# 
# 
# ggplot(data = data.frame(x = c(mu1 - s1*3, mu1 + s1*3)), aes(x = x)) +
#   stat_function(fun = dnorm, args = list(mean = mu1, sd = s1),
#                 color = "orange") +
#     geom_segment(x = shadeFrom, y = 0, xend = shadeFrom, yend = dnorm(shadeFrom, mu1, s1), color = "steelblue") +
#   theme_bw() +
#   labs(title = "Two normal curves")
```


- todo: fix exercise 2