---
title: "Hypothesis tests and confidence intervals"
subtitle: "STA 101"
format: html
---

## Bulletin

- Exam 2 Thursday
  - no lab Friday, no TA office hours Friday/Monday

## Today

By the end of today you will...

- define type I and type II error
- compare hypothesis tests with confidence intervals
- practice conducting hypothesis tests


## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae20.qmd",
destfile = "ae20.qmd")
```

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
library(openintro)
```



## Practice

### Load data

The `stent30` data set comes from the `openintro` package and is from a study conducted in 2011 on the effects of arterial stents as a therapy for stroke patients. See the original publication:

Chimowitz MI, Lynn MJ, Derdeyn CP, et al. 2011. Stenting versus Aggressive Med- ical Therapy for Intracranial Arterial Stenosis. New England Journal of Medicine 365:993- 1003. doi: 10.1056/NEJMoa1105335.

or check `?stent30` for more information.

```{r}
data(stent30)
```

```{r}
glimpse(stent30)
```

#### Exercise 1

Do stents affect stroke outcome in patients?

-   Write the null and alternative hypothesis. Report the observed statistic.

-   Simulate under the null and visualize the null distribution.

-   Compute and report the p-value, compare to $\alpha = 0.05$ and make a conclusion with appropriate context


## Confidence intervals and hypothesis tests

Here we revisit the data from the first three seasons of NC Courage games (2017-2019).

```{r message = F, warning = F}
courage = read_csv("https://sta101-fa22.netlify.app/static/labs/data/courage.csv")
```

```{r}
glimpse(courage)
```

Do National Women's Soccer League (NWSL) teams have a home-field advantage? We'll answer this question in a few separate ways.

Hypothesis testing framework: does NC Courage score a significantly different number of points (on average) away than at home?

#### Exercise 2 

- Create a new column `location` that tells you whether the courage are "home" or "away"
- Create a new column `pts` that always reports the Courage points scored in a game.
- Save your result as a new data frame titled `courage2`.

```{r}
# code here
```

#### Exercise 3

To answer the question *does NC Courage score a significantly different number of points (on average) away than at home?*

- Write the null and alternative hypothesis. Report the observed statistic.

- Simulate under the null and visualize the null distribution.

- Compute and report the p-value, compare to $\alpha = 0.05$ and make a conclusion with appropriate context


```{r}
# code here
```


#### Exercise 4

- Report the mean difference between away and home games and report a 95% bootstrap confidence interval. Use `set.seed(3)` and `reps=5000` Interpret your interval in context.

```{r}
# code here
```


#### Exercise 5

Is there a better way we could investigate whether or not the Courage have a home-field advantage? Why?

## Notes

### Type 1 and Type 2 Errors

| Truth         | Reject the null | Fail to reject the null |
|---------------|-----------------|-------------------------|
| $H_0$ is true | Type 1 error    | ✔️                      |
| $H_A$ is true | ✔️              | Type 2 error            |

The significance level, $\alpha$, is the probability of a type 1 error. In some contexts, a type 1 error may be referred to as a "false positive" and a type 2 error as a "false negative".

Intuitively, by considering extremes, one can see a trade-off exists between type 1 and type 2 error. \
\
If $\alpha = 0$, then the p-value stands no chance of being smaller than $\alpha$ and we always fail to reject the null. This makes type 1 errors impossible.\
\
Similarly, if $\alpha = 1$, then all p-values will be smaller than $\alpha$ and type 2 errors will become impossible, because we will always reject the null.

$\beta$ is used to denote the probability of a type 2 error.

The power of a test is $1 - \beta$, which is the probability that your test rejects the null hypothesis when the null hypothesis is false.

## Why it's important to be careful with interpretation

#### (And why hypothesis tests don't tell the whole story)

The data for this example comes from Confounding and Simpson’s paradox^[Julious, S A, and M A Mullee. “Confounding and Simpson’s paradox.” BMJ (Clinical research ed.) vol. 309,6967 (1994): 1480-1. doi:10.1136/bmj.309.6967.1480] by Julious and Mullee.

The data examines 901 individuals with diabetes and includes the following variables

- `insulin_dep`: whether or not the patient has insulin dependent or non-insulin dependent diabetes
- `age`: whether or not the individual is less than 40 years old
- `survival`: whether or not the individual survived the length of the study

```{r load-diabetes, warning = F, message = F}
diabetes = read_csv("https://sta101.github.io/static/appex/data/diabetes.csv")
```

Flex Aisher thinks people with insulin dependent diabetes actually survive longer than those without insulin dependence. Flex wants to formally test his hypothesis.

Let $p_{d}$ be the probability of insulin *dependent* survival and $p_{i}$ be the probability of insulin independent survival.

$$
H_0: p_{d} - p_{i} = 0
$$

$$
H_A: p_{d} - p_{i} > 0
$$

At first glance the data seem to back up his claim...

#### Exercise 6

Compute the probability of survival and death for diabetic individuals with and without insulin dependence.

```{r}
#  code here
```

#### Exercise 7

Is Flex's claim significant at the $\alpha = 0.05$ level? Perform a hypothesis test and report your results.

```{r}
# code here
```

#### Exercise 8

Is the aggregate data misleading? Use the code chunk below to investigate further.

```{r}
# code here
```

