---
title: "Regression + Inference"
subtitle: "STA 101"
format: html
---

## Bulletin

- Draft final project report due Friday December 2
- This Friday is last lab before peer-review in two weeks

## Today

By the end of today you will...

- understand regression hypothesis testing
- interpret p-values in a regression framework

## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae22.qmd",
destfile = "ae22.qmd")
```

## Load packages

```{r load-libraries, message = F, warning = F}
library(tidyverse)
library(tidymodels)
library(viridis)
```

## Load data

```{r load-data, warning = F, message = F}
Seoul_Bikes = read_csv("https://sta101.github.io/static/appex/data/Seoul_Bikes.csv") 
Seoul_Calendar = read_csv("https://sta101.github.io/static/appex/data/Seoul_Calendar.csv")

bikes = left_join(Seoul_Bikes, Seoul_Calendar)
```

This data set was originally analyzed in two studies^[Sathishkumar V E, Jangwoo Park, and Yongyun Cho. 'Using data mining techniques for bike sharing demand prediction in metropolitan city.' Computer Communications, Vol.153, pp.353-366, March, 2020; Sathishkumar V E and Yongyun Cho. 'A rule-based model for Seoul Bike sharing demand prediction using weather data' European Journal of Remote Sensing, pp. 1-18, Feb, 2020] of predicting bike-rental usage in Seoul, South Korea. For this lecture, the data was sourced from [UCI Machine Learning Repository](https://archive.ics.uci.edu).

Code book:

- `Date`: the date
- `rented_bikes`: total number of bikes rented on a given day
- `temp_c`: mean daily temperature (Celsius)
-  `humidity_pct`: mean daily humidity
- `wind_speed`: mean daily windspeed
- `snowfall_cm`: mean daily snowfall (in cm)
- `season`: the season
- `holiday`: whether or not the day is a holiday

```{r glimpse-data}
glimpse(bikes)
```

## Notes

### Hypothesis testing in a regression framework


```{r plot-1}
bikes %>%
  ggplot(aes(x = temp_c, y = rented_bikes, color = holiday)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(x = "Temperature", y = "Rented No. Bikes", color = "Holiday?",
       title = "Rented Bicycles in Seoul") +
  scale_color_manual(values = c("red", "#00539B")) +
  geom_smooth(method = 'lm')
```

#### Exercise 1

Write the full model equation to match the figure above to predict the number of bikes rented on a given day based on the temperature outside and whether or not the day is a holiday.


$$
\text{Your equation here}
$$


#### Exercise 2

Fit the model above and examine the tidied output. What are the p-values associated with each predictor?

```{r}
# code here
```

The p-value output is associated with a typical hypothesis test... but what's the null and alternative?

The main idea is that if a predictor (e.g. holiday) does not help us explain bike rental numbers then its associated $\beta$ might as well be 0. Within the framework of hypothesis testing:

$H_0$: $\beta_{\text{holiday}} = 0$

$H_A:$ $\beta_{\text{holiday}} \neq 0$

For OLS regression, our test statistic is 

$$
T = \frac{\hat{\beta} - 0}{\text{SE}_{\hat{\beta}}} \sim t_{n - 4}
$$
We want to see if our observed statistic, $T$, falls far in the tail under the null.

`R` takes care of much of this behind the scenes with the tidy output and reports a p-value for each $\beta$ by default.

#### Exercise 3

Calculate the p-value associated with $\beta_{holiday}$ manually using the equation above. Note: in a regression setting, the degrees of freedom is the number of observations minus the number of $\beta$s.

Compare the p-value to one reported in the tidy output.

```{r manual-p-value}
# code here
```

- Is $\beta_{\text{holiday}}$ significant at the $\alpha = 0.05$ level? State your conclusion.

- Looking at the tidied output, are any of the $\beta$s not significant at the $\alpha = 0.05$ level?

- Change the model from an interaction effects to a main effects model. What do you notice?

