---
title: "Linear regression III"
subtitle: "STA 101"
format: html
---


## Bulletin

- Lab 3 due tonight
- Looking towards next week, please fill out [this optional form](https://forms.gle/rLbDjjiemgRSctmp9) to request group members (from your lab) to work on the projects with.
- [Exam 1](https://sta101-fa22.netlify.app/static/exams/exam01.html) released tonight and due Monday
  - no TA office hours Friday/Monday
  - ask questions early

## Warm-up

Check you understanding! Answer the following...

1. To "fit" a linear model means...[fill in the blank]

2. Is $y = \beta_0 + \beta_1 \log(x_1)+ \beta_2 x_2^2 + \epsilon$ a linear model? Why or why not?

## Today

By the end of today you will...

- fit and interpret models with categorical predictors
- fit and interpret models with interactive effects

## Getting started

Download this application exercise by pasting the code below into your **console** (bottom left of screen)

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae8.qmd",
destfile = "ae8.qmd")
```


## Load packages and data

```{r load-libraroies, warning = F, message = F}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
```

Today we will return to our Palmer penguins data set

```{r load-penguins, warning = F, message = F}
data(penguins)
```

Use `?penguins` or [click here](https://github.com/allisonhorst/palmerpenguins) for more info.

## Notes

## Main effects

Up until now, we've seen models that look like this:

$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \ldots + \epsilon
$$

Here's an example:

$y$: body mass (g)

$x_1$: bill length (mm)

$x_2$: 1 if island Dream, 0 otherwise

$x_3$: 1 if island Torgersen, 0 otherwise

$$
y = \beta_0 + {\beta_1} x_1 + {\beta_2} x_2 + {\beta_3} x_3 + \epsilon
$$

Notice that 

- We have a categorical predictor `island` that takes three values: Dream, Torgersen, and Biscoe.

- Despite taking three values, there are only two island variables in the model. One for `Dream`  and one for `Torgersen`. Biscoe island is considered the default. This always occurs when we have a categorical variable -- one category is considered the default.

- Bill length only impacts body mass via the term $\beta_1 x_1$. 

- $x_2$ and $x_3$ can be thought of as turning on or off a constant.

Let's visualize the main effects model below.

```{r main, echo = F, message = F, warning = F}
# main_fit = linear_reg() %>%
#   set_engine("lm") %>%
#   fit(body_mass_g ~ bill_length_mm + island, data = penguins)
#   main_fit %>%
#   tidy()

b0 = 1225.79107
bD = -919.07448
bT = -523.29201
bBill = 77.11933

# show_col(hue_pal()(3))  

penguins %>%
ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point() +
  theme_bw() +
  geom_abline(intercept = b0, slope = 77.11933, color = "#F8766D", lwd = 1) +
  geom_abline(intercept = b0 + bD, slope = 77.11933, color = "#00BA38", lwd = 1) +
  geom_abline(intercept = b0 + bT, slope = 77.11933, color = "#619CFF", lwd = 1) +
  labs(x = "Bill length (mm)",
       y = "Body mass (g)",
       title = "Island main effect")
```


We can fit the "main effects" model above with our standard procedure:

```{r main-effects-1}
main_fit = linear_reg() %>%
  set_engine("lm") %>%
  fit(body_mass_g ~ bill_length_mm + island, data = penguins)

  main_fit %>%
  tidy()
```

If we want to know how bill length relates to body mass for penguins on island Biscoe, we plug in $0$ for $x_2$ and $x_3$ and write the resulting model. If we repeat as appropriate for each island, the result is 3 separate fitted models:

Biscoe:

$$
\hat{y} = 1225.8 + 77.1 x_1 
$$

Dream:

$$
\hat{y} = 1225.8 + 77.1 x_1 - 919.1
$$


Torgersen:

$$
\hat{y} = 1225.8 + 77.1 x_1 -523.3
$$

Notice that in each case, the slope associated with bill length ($x_1$) is the same.

### Interaction effects

Interaction effect models contain products of predictors, e.g.

$$
y = {\beta_0} + {\beta_1} x_1 + {\beta_2} x_2 + {\beta_3} x_3 +  {\beta_4} x_1 x_2 + {\beta_5} x_1 x_3 + \epsilon
$$

Here we have an interaction between bill length and island ($\beta_4 x_1 x_2$ and $\beta_5 x_1 x_3$).

Take-away idea: $x_1$ is related to $y$ but the relationship changes depending on $x_2$ and $x_3$.

The simplest scenario is one of "group membership". In other words, knowing the group your measurement belongs to affects the relationship between $x_1$ and $y$.

Here, we see bill length ($x_1$) show up multiple times in our linear model paired with islands. In other words, the relationship between bill length and body mass depends on the island a penguin is from.

We fit this interaction model using the code below:

```{r interaction-fit}
interaction_fit = linear_reg() %>%
  set_engine("lm") %>%
  fit(body_mass_g ~ bill_length_mm * island, data = penguins)

  interaction_fit %>%
  tidy()
```


```{r interaction, echo = F, warning = F, message = F}
penguins %>%
ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
geom_point() +
theme_bw() +
geom_smooth(method = 'lm', se = F) +
labs(x = "Bill length (mm)",
     y = "Body mass (g)",
     title = "Island interaction effect")
```

Interpreting interactions can be difficult, especially without writing things down. To make it easier, we will compare the implied linear models:

Plug in 0 for `islandDream` ($x_2$) and 0 for `islandTorgersen` ($x_3$) to get the linear model for `islandBiscoe` penguins

Plug in 1 for `islandDream` ($x_2$) and 0 for `islandTorgersen` ($x_3$) to get the linear model for `islandDream` penguins

Plug in 0 for `islandDream` ($x_2$) and 1 for `islandTorgersen` ($x_3$) to get the linear model for `islandTorgersen` penguins

- Biscoe fitted model:

$$
\hat{y} = -1726.0+ 142.3 x_1
$$

- Dream fitted model:

$$
\hat{y} = -1726.0 + 142.3 x_1 + 4478.7 -120.6 x_1
$$

Combine terms:

$$
\hat{y} = 2752.7 + 21.7 x_1 
$$

#### Exercise 1 

Write out the fitted model for Torgersen island below.

- Torgersen model:
$$
\hat{y} = [\text{write here}]
$$

### Interpreting

Now we can interpret the interaction model by comparing bill length slopes between islands.

- For a unit increase in bill length of a penguin from the island Dream, how much do we expect the body mass to increase?

#### Exercise 2

- You measured the bill length of a penguin from island Biscoe and a penguin from island Torgersen a year ago. You re-measure them today and find the bill length of each one grew by exactly 2 mm. How much mass do you expect each penguin to have gained?

#### Exercise 3

Are the intercepts meaningful?


#### Exercise 4

Is the relationship between Body mass (g) and Bill depth (mm) positive or negative? Create a convincing argument from the data.

#### Exercise 5

Build a linear model of body mass using bill depth and one other predictor of your choosing (hint: see previous exercise!)

- Write out a linear model with both predictors and fit the model.

- Fit the linear model

- Do you prefer this model to the interaction effects model from a previous exercise? Why?



