---
title: "Probability prepare"
subtitle: "STA 101"
format: html
---

## Getting started

Download this prepare file by pasting the code below into your **console** (bottom left of screen)

    download.file("https://sta101-fa22.netlify.app/static/appex/prepareProbability.qmd",
      destfile = "prepareProbability.qmd")

## Goals

- be able to define and give examples of sample space, outcomes, events, probabilities, conditional, marginal, joint and independent probabilities


## Load packages

```{r load-libraries, message = F, warning = F}
library(tidyverse)
library(fivethirtyeight)
```

## Notes

### Sample space

The **sample space** is the set of all possible outcomes of an experiment.

#### Discrete examples

- Experiment 1: You flip a coin once. The sample space is $\{H, T\}$.

We separate each outcome by a comma and use brackets $\{ \}$ to denote a "set".

- Experiment 2: You flip a coin twice. The sample space is $\{ HH, HT, TH, TT\}$

- Experiment 3: You roll a die once. The sample space is $\{ 1, 2, 3, 4, 5, 6\}$

- Experiment 4: You send out a survey asking participants whether they prefer cats or dogs. The sample space is $\{ \text{Cats}, \text{Dogs} \}$

- Experiment 5: A car manufacturer makes 100 vehicles. You count the number of recalls. The sample space is $\{0, 1, 2, 3, \ldots, 99, 100 \}$

#### Continuous examples

- Experiment 6: You observe the numeric grade you earn in a course. The sample space is $[0, 100]$

Here we write the lower bound and upper bound of the sample space and assume we can observe all values in-between. Brackets, $[$ $]$, are inclusive of the end values while parentheses, $($ $)$, are not.

- Experiment 7: You measure the tail length of American alligators The sample space is $(0, c ]$ feet where $c$ is the maximum tail length of an alligator, e.g. $c$ might be approximately 10.

- Experiment 8: You measure the geographic coordinates (longitude and latitude) of a COVID case. The sample space is $[-90, 90]$ for latitude and $[-180, 180 ]$ for longitude. 

### Events

An **event** is a collection of 1 or more outcomes. Two events are said to be **disjoint** if they cannot occur at the same time.

#### Examples

- You roll a die once. Let *A* be the event that you roll an even number, i.e. *A* $= \{2, 4, 6 \}$. Let *B* be the event you roll a 1 or a 2, i.e. *B* $= \{1, 2 \}$. *A* and *B* are **not** disjoint.

- A car manufacturer makes 100 vehicles. You count the number of recalls. Let *C* be the event you see fewer than 10 recalls. *C* $= \{0, 1, 2, 3, \ldots, 8, 9 \}$

- You observe the numeric grade you earn in a course. Let *D* be the event you receive a letter grade of "A". *D* $= [93, 100]$. Let *E* be the event that you earn a "B" or worse. *E* $= [0, 87)$. *D* and *E* **are disjoint** events because they cannot occur simultaneously.


### Random variables 

Random variables are functions that map outcomes to numbers. An **indicator random variable** takes values *1* and *0* to indicate whether or not an event occurs.

```{r}
data(bob_ross) # within fivethirtyeight package
bob_ross %>%
  head(10)
```


### Probability

A **probability** is the long-run frequency of an *event*. In other words, the proportion of times we would see an event occur if we could repeat an experiment an infinite number of times. Probabilities take values between 0 and 1 inclusive.

If *A* and *B* are two disjoint events, then the probability of *A* or *B* occurring is equal to the probability of *A* plus the probability of *B*. More concisely, Pr(*A* or *B*) = Pr(*A*) + Pr(*B*).

#### More definitions

Let *A* and *B* be two events.

- Marginal probability: The probability an event occurs regardless of values of the other event
  - P(*A*)
  - Example: What's the probability that, in a randomly selected episode of Bob Ross, the painting features clouds?
- Joint probability: The probability two or more events simultaneously occur
  - Example: What's the probability that, in a randomly selected episode of Bob Ross, the painting features clouds and mountains?
  - P(*A* and *B*)
- Conditional probability: The probability an event occurs given the other has occurred
  - P(*A*|*B*) or P(*B*|*A*)
  - Example: What is the probability that a Bob Ross painting features clouds in season 1?
  - P(*A*|*B*) = P(*A* and *B*) / P(*B*)
- Independent events: Knowing one event has occurred does not lead to any change in the probability we assign to another event.
  - P(*A*| *B*) = P(*A*) or P(*B*|*A*) = P(*B*)
  - Example: P(lakes | rivers) = P(lakes)

