---
title: "Peer review instructions"
subtitle: "STA 101"
date: "December 2, 2022"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer-2.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Peer-review upload

- Go to the [course sakai page](https://sakai.duke.edu/) and click discussions on the left-hand side navigate to `Discussions` > `Peer-review` > `Start a new conversation`

- Make the title of your post "Team name -- project name" as in the example. Upload the most current draft  of your project report as a PDF.

```{r echo = F}
countdown::countdown(minutes = 5, seconds = 0)
```

---

## Peer-review

- Check the [team assignments page](https://duke.box.com/s/gigjvdrogsh8nur2tag78n6gmb4ey0hk) and click on your lab tab at the bottom of the sheet. Next find `peer-review` column at the top to access peer-review assignments. 

- On the discussion forum, navigate to your assigned group's "Conversation" and click `Reply to initial message`. Answer the questions on the following slides and click `Post`. It might help to assign 1 group member to each slide of questions.

---

## Project motivation

1. Is the research question clear and concisely stated? Be specific.

2. Is the topic well motivated?

3. Is enough context provided, or is anything unclear? E.g. variable names etc.

4. Are the observations and relevant variables described?

---

## Methodology

1. Do the methods proposed help answer the research question?

2. Are any methods (regression, hypothesis testing etc.) adequately introduced and defended? E.g. if using regression, the full model equation should be listed here.

3. Are the variables used pertinent to the question at hand? Are strengths and weaknesses of variables discussed?

4. Include any additional comments, questions, or suggestions related to the choice of variables, statistical models, or general methods


---

## Cohesiveness of results

1. Do the results match the methods? 

2. If statistical methods are used, are relevant fitted equations and/or p-values, confidence intervals etc. reported?

3. Do any included figures/plots help tell a story related to the initial research question? Are the plots useful? Note if any plots are unclear.

4. Is the research question answered accurately and effectively? Is the answer to the research question summarized and supported by statistical arguments or effective visualizations?

---

## Formatting

1. Is the paper professionally presented and generally free of distracting errors or other issues, including (but not limited to) insufficient organization or formatting; poor grammar, spelling, or punctuation; or too-small font? Is the overall paper easily readable for someone with your expert level of knowledge? Note any concerns here.

2. Are all figures legible and appropriately labeled?

3. Is the author name and team name printed at the top of the document?

4. What questions and/or general feedback do you have for the authors?
