---
title: "CLT Continued and Intro to Hypothesis Testing"
output: html_document
date: "July 21, 2022"
---


## Bulletin

- Exam 2 due Monday July 25 at 11:59pm
- No lab today

## Today

By the end of today you will...

- Construct a CLT CI
- Be familiar with the terms "p-value", "null-distribution", "null hypothesis", "alternative hypothesis"
- Compute a p-value


## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101.github.io/static/appex/ae11.Rmd",
destfile = "ae11.rmd")
```

## Load packages and data

```{r load-libraries, message = F, warning = F}
library(tidyverse)
library(tidymodels)
```


```{r load-data, message = F, warning = F}
pokemon = read_csv("https://sta101.github.io/static/appex/data/pokemon.csv")
```


## Notes

### The standard normal

The standard normal distribution is a normal distribution with mean 0 and standard deviation 1. Often, the character "Z" is used to denote a standard normal random variable.

$$
Z \sim N(0,1)
$$


```{r standard-normal}
ggplot(data = data.frame(x = c(0 - 1*3, 0 + 1*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1),
                color = "black") +
  labs(x = "Z", y = "Density", title = "The standard normal") +
  theme_bw()
```

We can write any normal distribution as a linear combination of a standard normal. For example,

If $X \sim N(10,4)$, then $X = 4Z + 10$. *Note that the second term inside of the parentheses is the standard deviation. In some textbooks the second term is variance.

Example:
We draw 1000 samples of x (gold) and 1000 samples of z (blue). If we multiply each z by the standard deviation (4) and add the mean (10), we equivalently obtain samples from x; this is shown in red.

```{r ex-linear-combo, warning = F, message = F}
set.seed(1)
normal_df = data.frame(z = rnorm(1000, mean = 0, sd = 1),
                 x = rnorm(1000, mean = 10, sd = 4))

sample_mean = normal_df %>%
  summarize(sample_mean = mean(x)) %>%
  pull()

normal_df %>%
  ggplot(aes(x = x)) +
  geom_histogram(aes(y = ..density..), fill = 'gold3', alpha = 0.5) +
  geom_histogram(aes(x = z, y = ..density..), fill = 'steelblue', alpha = 0.7) +
  # geom_density(aes(x = z), color = 'steelblue') +
  geom_density(aes(x = z*4 + 10), color = 'red') + 
  theme_bw() +
  labs(x = "", y = "", title = "Sampling from two normals")

```

Similarly, we can find confidence intervals using either the true distribution or by transforming Z. Our `sample_mean` (computed above) is `r sample_mean`. We can look at the associated 95% confidence interval:


```{r}
# lower bound
qnorm(0.025, mean = sample_mean, sd = 4)
qnorm(0.025, mean = 0, sd = 1) * 4 + sample_mean

# upper bound
qnorm(0.975, mean = sample_mean, sd = 4)
qnorm(0.975, mean = 0, sd = 1) * 4 + sample_mean
```

More generally, we construct a confidence interval

$$
\bar{x} \pm z^* \times \frac{\sigma}{\sqrt{n}}
$$

where $z^*$ denotes the z-score associated with the quantiles of interest for a given confidence level.

### Practical confidence intervals

We don't know the true population mean $\mu$ and standard deviation $\sigma$, how do we use CLT to construct a confidence interval?

We approximate $\mu$ by $\bar{x}$ and $\sigma$ by $s$. However $s$ may be smaller than $\sigma$ and our confidence interval could be too narrow, for example, run the code below:

```{r ex-sd}
set.seed(6)
samples = rnorm(10, mean = 0, sd = 1)
sd(samples)
```

This was just for 1 random seed. If you repeat with no seed, you will find that sometimes $s$ is above and sometimes $s$ is below the true standard deviation.

To account for this uncertainty, we will use a distribution with thicker tails. This sampling distribution is called a **t-distribution**.

```{r t-versus-normal}
ggplot(data = data.frame(x = c(0 - 1*3, 0 + 1*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1),
                color = "black") +
  stat_function(fun = dt, args = list(df = 3),
                color = "red",lty = 2) + theme_bw() +
  labs(title = "Black solid line = normal, Red dotted line = t-distribution")
```

The t-distribution has a bell shape but the extra thick tails help us correct for the variability introduced by using $s$ instead of $\sigma$.

The t-distribution is always centered at zero and has a single parameter: degrees of freedom. The degrees of freedom describes the precise form of the bell-shaped t-distribution. In general, we'll use a t-distribution with $df=n−1$ to model the sample mean when the sample size is $n$.

We can use `qt` and `pt` to find quantiles and probabilities respectively under the t-distribution.

### Confidence interval
To construct our 
practical confidence interval (where we don't know $\sigma$) we use the t-distribution:

$$
\bar{x} \pm t^*_{n-1} \times \frac{s}{\sqrt{n}}
$$

#### Exercise 1

- Calculate the 95% confidence interval for pokemon height using the t-distribution.

```{r clt-ci}
# code here
```

How does this compare to a 95% bootstrap confidence interval?

```{r bootstrap-ci}
# code here
```


---

## Introduction to hypothesis testing


### Is this a fair coin?

We'll record 1 if the coin is "Heads" and 0 if the coin lands "Tails".

```{r fair-coin}
# coin_flips = c()
```

If the coin is fair, what is the probability of seeing the outcome we saw? To answer this question we'll setup a statistical model:

$$
\text{# heads} \sim Binomial(n, p)
$$

where $p$ is the probability of a heads and $n$ is the total number of coin flips. 

#### Exercise 2

- If the coin is fair, what would $p$ be?

- Using `R`, flip a fair coin 6 times and count the number of heads. Next, repeat this experiment 1000 times and count the proportion of times you observe 6 heads.

```{r compute-probability}
# code here
```

- What is the probability of observing 6 heads in 6 coin flips?


## Hypothesis testing framework

You may not have realized it but you just performed a hypothesis test!

You setup a null hypothesis: $H_0$. The null hypothesis is a hypothesis you set up and then try and knock down. Conceptually, it's the "nothing special is going on" hypothesis. Formally, the null hypothesis makes a **claim** or **assumption** about a population parameter. 

In this case, it was the assumption that the coin is fair. Mathematically, we write this:

$$
H_0: p = 0.5
$$

Conceptually, $p$ is the probability of flipping a head if we flipped the coin infinitely many times.

Since we computed the probability of observing **all heads**, we were fundamentally interested in if $p > 0.5$. This is our alternative hypothesis, $H_A$. In mathematical notation, we write

$$
H_A: p > 0.5
$$

Next, we **simulated under the null**. This means that we simulated what the coin flips would have looked like **if the null was true**. In this context, this means we simulated as **if the coin was fair**.

Finally, we check to see where our actual observed data places under the null distribution. If it's **way out in the tail**, we reject the null. If its not way out in the tail, we **fail to reject the null**.

How can we make "way out in the tail" more precise? Well, it's arbitrary and context-dependent. In some contexts it is popular to use a cut-off of $0.05$.

#### Exercise 3
 
Assume we continue flipping our coin for a total of 30 coin flips and observe 26 heads and 4 tails. What is the probability of seeing 26 **or more** heads if the coin is fair?

```{r ex-3}
# code here
```


## p-values

You might not realize it, but you just computed a p-value... again!

A p-value is a probability. It's the **tail** probability associated with your alternative hypothesis.

The alternative hypothesis must always relate to the null. Here we had three options:

- $H_A: p < 0.5$, the coin is biased to land tails
- $H_A: p > 0.5$, the coin is biased to land heads
- $H_A: p \neq 0.5$, the coin is biased

Let's look offline at what each one would like here.

#### Exercise 4

Compute the p-value associated with each of the alternative hypotheses above.

```{r}
# code here
```

Make a conclusion based on a significance level of 0.05. In other words, 

- if p < 0.05, reject the null.
- if p > 0.05, we fail to reject the null.

As you can see our conclusion depends on our alternative hypothesis. For this reason, it is important to set up an alternative hypothesis **before** looking at the data.
