---
title: "The bootstrap"
subtitle: "STA 101"
format: html
---

## Bulletin

- Lab 05 due tonight

## Today

By the end of today you will...

- be able to draw a bootstrap sample and calculate a bootstrap statistic 
- use `infer` to obtain a bootstrap distribution 
- calculate a confidence interval from the bootstrap distribution
- interpret a confidence interval in context of the data 

## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae15.qmd",
destfile = "ae15.qmd")
```

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

## Load data

```{r load-data, warning = F, message = F}
manhattan = read_csv(
  "https://sta101-fa22.netlify.app/static/appex/data/manhattan.csv"
  )
```

## Notes (for reference)

**Bootstrapping** is a re-sampling technique. The key idea is you have already collected a **sample** of size $N$ from the **population**. To create a **bootstrap** sample, you sample *with replacement* from your original sample $N$ times.

Let's say you measure the height of five Duke students in meters:

```{r}
heights = c(1.51, 1.62, 1.89, 2.01, 1.78)

students = data.frame(heights)
```

There are many ways to create a bootstrap sample in R. We will focus on the `tidy` way below. which uses the `infer` package that loads with `tidymodels`.

#### Example

```{r method-2}
set.seed(2)
students %>%
  specify(response = heights) %>%
  generate(reps = 1, type = "bootstrap")
```

::: {.callout-note appearance="simple"}
## Note

Sampling is **random**. Notice the *seed* above ensures we get the same bootstrap sample.
:::

From here, we can compute a bootstrap statistic. E.g.

```{r statistic-computation}
set.seed(2)
students %>%
  specify(response = heights) %>%
  generate(reps = 1, type = "bootstrap") %>%
  calculate(stat = "median")

```


## Example: rent in Manhattan

On a given day in 2018, twenty one-bedroom apartments were randomly selected on Craigslist Manhattan from apartments listed as "by owner". The data are in the `manhattan` data frame.  We will use this sample to conduct inference on the typical rent of 1 bedroom apartments in Manhattan. 

### Part 1: Drawing a bootstrap sample

Let's start by using bootstrapping to estimate the **mean** rent of one-bedroom apartments in Manhattan. 

#### Exercise 1

What is a **point estimate** (i.e. single number summary) of the typical rent? 

#### Exercise 2

Let's bootstrap! 

- To bootstrap we will sample with replacement by drawing a value from the box. 
- How many draws do we need for our bootstrap sample? 

Fill in  the values from the bootstrap sample conducted in class. Once the values are filled in, un-comment the code. 

```{r boot-sample}
# class_bootstrap = c()

```

#### Exercise 3

- About what value do you expect the bootstrap statistic to take? 
- Calculate the statistic from the bootstrap sample. 

```{r calc-boot-stat}
# add code

```

### Part 2: Bootstrap confidence interval

**We will calculate a 95% confidence interval for the mean rent of one-bedroom apartments in Manhattan.**

We start by setting a seed to ensure our analysis is reproducible. 

#### Generating the bootstrap distribution

We can use R to take many bootstrap samples, compute a statistic and then view the bootstrap distribution of that statistic.

Un-comment the lines and fill in the blanks to create the bootstrap distribution of sample means and save the results in the data frame `boot_dist`.    

Use **1000 reps** for the in-class activity. (You will use about 10,000 reps for assignments outside of class.)

```{r create-bootstrap-dist}
set.seed(7182022)

boot_dist = manhattan #%>%
  #specify(______) %>%
  #generate(______) %>%
  #calculate(______)
```

- How many rows are in `boot_dist`? 
- What does each row represent? 
- What are the variables in `boot_dist`? What do they mean?

#### Visualize the bootstrap distribution 

A **sample statistic is a random variable**, we can look at its distribution.

Visualize the bootstrap distribution using a histogram. Describe the shape and center of the distribution.

```{r viz-bootstrap}
# add code

```

#### Calculate the confidence interval

Uncomment the lines and fill in the blanks to construct the 95% bootstrap confidence interval for the mean rent of one-bedroom apartments in Manhattan. 

```{r confidence-interval-95}
#___ %>%
#  summarize(lower = quantile(______),
  #          upper = quantile(______))
```

#### Interpret the interval

Write the interpretation for the interval calculated above. 

- Question: Does a confidence interval have to be symmetric?

- What is one advantage to using a 90% confidence interval instead of a 95% confidence interval to estimate a parameter? - What is one advantage to using a 99% confidence interval instead of a 95% confidence interval to estimate a parameter? 
