---
title: "CLT and Confidence Intervals"
subtitle: "STA 101"
format: html
---

## Bulletin

- Lab 6 due tonight
- Project proposal due Friday (tomorrow)

## Today

By the end of today you will...

- Use CLT to construct confidence intervals

## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae17.qmd",
destfile = "ae17.qmd")
```

## Load packages

```{r load-libraries, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

## Load Data: Pokemon

We will be using the `pokemon` data set, which contains information about 42 randomly selected Pokemon (from all generations). You may load in the data set with the following code:

```{r load-data-2, warning = FALSE, message = FALSE}
pokemon = read_csv("https://sta101.github.io/static/appex/data/pokemon.csv")
```

In this analysis, we will use CLT-based inference to draw conclusions about the mean height among all Pokemon species.

#### Exercise 1

Let's start by looking at the distribution of `height_m`, the typical height in meters for a Pokemon species, using a visualization and summary statistics.

```{r height-histogram}
ggplot(data = pokemon, aes(x = height_m)) +
  geom_histogram(binwidth = 0.25, fill = "steelblue", color = "black") + 
  labs(x = "Height (in meters)", 
       y = "Distributon of Pokemon heights")
```

```{r height-stats}
pokemon %>%
  summarise(mean_height = mean(height_m), 
            sd_height = sd(height_m), 
            n_pokemon = n())
```

In the previous lecture we were given the mean, $\mu$, and standard deviation, $\sigma$, of the population. That is unrealistic in practice (if we knew $\mu$ and $\sigma$, we wouldn't need to do statistical inference!).

Today we will start on using the Central Limit Theorem to draw conclusions about the $\mu$, the mean height in the population of Pokemon.

-   What is the point estimate for $\mu$, i.e., the "best guess" for the mean height of all Pokemon?

-   What is the point estimate for $\sigma$, i.e., the "best guess" for the standard deviation of the distribution of Pokemon heights?

#### Exercise 2

Before moving forward, let's check the conditions required to apply the Central Limit Theorem. Are the following conditions met:

-   Independence?
-   Sample size/distribution?


#### Central limit theorem

Remember, when the independence and sample size assumptions are met, the central limit theorem states

$$
\bar{x} \sim N(\mu, \sigma / \sqrt{n})
$$
If we know $\sigma$, we can construct a symmetric confidence interval for the true mean easily using `qnorm()`.

For example, if the true standard deviation in pokemon height is 0.4 meters, then to construct a 95% confidence interval:

```{r}
xbar = pokemon %>%
  summarize(xbar = mean(height_m)) %>%
  pull(xbar)

qnorm(c(0.025, 0.975), mean = xbar, sd = 0.4)
```

This can be equivalently expressed

```{r}
zscore = qnorm(0.025)
xbar + zscore*0.4
xbar - zscore*0.4
```

where we use the fact that we can write any normal distribution as a linear combination of a standard normal. For example,

if $X \sim N(0.928, .4)$, then $X = .4Z + 0.928$ where $Z$ is standard normal, in other words $Z \sim N(0, 1)$.

In general, the confidence interval can be written as 

$$
\bar{x} \pm z^* \times \sigma
$$
where $z^*$ is the quantile of a standard normal distribution associated with our level of confidence.

What about when we don't know $\sigma$?

### Practical confidence intervals

We don't know the true population mean $\mu$ and standard deviation $\sigma$, how do we use CLT to construct a confidence interval?

We approximate $\mu$ by $\bar{x}$ and $\sigma$ by the same standard deviation $s$. However $s$ may be smaller than $\sigma$ and our confidence interval could be too narrow, for example, run the code below to compute the standard deviation of three draws from a standard normal.

```{r ex-sd}
set.seed(6)
samples = rnorm(3, mean = 0, sd = 1)
sd(samples)
```

This was just for 1 random seed. If you remove the seed and repeat the simulation, you will find that $s$ is sometimes above and sometimes below the true standard deviation.

To account for this uncertainty, we will use a distribution with thicker tails. This sampling distribution is called a **t-distribution**.

```{r t-versus-normal}
ggplot(data = data.frame(x = c(0 - 1*3, 0 + 1*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1),
                color = "black") +
  stat_function(fun = dt, args = list(df = 3),
                color = "red",lty = 2) + theme_bw() +
  labs(title = "Black solid line = normal, Red dotted line = t-distribution")
```

The t-distribution has a bell shape but the extra thick tails help us correct for the variability introduced by using $s$ instead of $\sigma$.

The t-distribution, like the standard normal, is always centered at zero. Therefore, the t-distribution has only a single parameter: degrees of freedom. The degrees of freedom describes the precise form of the bell-shaped t-distribution. In general, we'll use a t-distribution with $df=n−1$ to model the sample mean when the sample size is $n$.

We can use `qt` and `pt` to find quantiles and probabilities respectively under the t-distribution.

### Confidence interval
To construct our 
practical confidence interval (where we don't know $\sigma$) we use the t-distribution:

$$
\bar{x} \pm t^*_{n-1} \times \frac{s}{\sqrt{n}}
$$

#### Exercise 3

- Calculate the 95% confidence interval for pokemon height using the t-distribution.

```{r clt-ci}
# code here
```

How does this compare to a 95% bootstrap confidence interval?

```{r bootstrap-ci}
# code here
```
