---
title: "Exploratory data I: penguins"
subtitle: "STA 101"
output: html_document
date: "June 28, 2022"
---

## Bulletin

- first lab today, due Friday at 11:59pm on Gradescope
- be sure to complete prepare material (on the [schedule](https://sta101.github.io/static/schedule/schedule.html)) before each class

## Today

By the end of today you will...

- gain familiarity with the 'big seven' functions of `dplyr`
- define and compute various statistics
- begin to gain familiarity with `ggplot`


## Getting started

Download this application exercise by pasting the code below into your **console** (bottom left of screen)

```{r download-rmd, eval=F}
download.file("https://sta101.github.io/static/appex/ae2.Rmd",
destfile = "ae2.rmd")
```

- A note on YAML. Create a new `.rmd`

## Load packages and data

```{r load-libraries, message=F, warning=F}
library(tidyverse)
library(palmerpenguins)
```

Type `?palmerpenguins` to learn more about this package. Or better yet, check it out [here](https://allisonhorst.github.io/palmerpenguins/).

```{r load-data, message=F, warning=F}
data(penguins)
```

#### Exercise 1: 
Look at the data, how many observations are there? How many variables?

## A package within a package...

When we load the `tidyverse` library, `dplyr` is packaged with it.

[dplyr](https://dplyr.tidyverse.org/), a grammar of data manipulation offers intuitive 'verb' functions that describe actions we commonly want to perform with data. The big 7 we'll cover today are:

- `mutate()` adds new variables that are functions of existing variables
- `select()` picks variables based on their names.
- `filter()` picks cases based on their values.
- `group_by()` sets us up to summarize across groups
- `summarize()` reduces multiple values down to a single summary.
- `arrange()` changes the ordering of the rows.
- `slice()` select, remove and duplicate rows based on their index

(as described at [https://dplyr.tidyverse.org/](https://dplyr.tidyverse.org/))

- Check out documentation with `?`

### Mutate

Approximate bill area (in $mm^2$) as bill length * bill depth:

```{r approx-bill-area}
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm)
```

### Select

It's hard to see bill length, depth and area in the same output, let's select a smaller subset of the variables to look at.

```{r remove-year}
# Example 1
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm) %>%
  select(-year)

# Example 2
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm) %>%
  select(species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, bill_area_mm2)
```

- A note on pipes `%>%` and a note on style.

## Filter

Let's just examine penguins on Dream island

```{r just-dream, eval=F}
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm) %>%
  select(-year)
# code here
```


## Group by + Summarize


#### Exercise 2:

Find mean bill area across sex. Fill in the blanks
```{r bill-area-across-sex, eval=F}
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm) %>%
  select(-year) %>%
  # filter for Dream
  group_by(___) %>%
  summarize(mean_bill_area_mm2 = ___)
```


## Arrange + Slice

Let's use `arrange()` and `slice()` to report the five penguins with the greatest bill area. 

```{r splitting-the-bill}
penguins %>%
  mutate(bill_area_mm2 = bill_length_mm * bill_depth_mm) %>%
  select(bill_area_mm2, bill_length_mm) %>%
  arrange(desc(bill_area_mm2))
```

#### Exercise 3:

Are these the same five penguins with the longest bills?

Optional hint: if you want to be exactly precise about which penguins are which, you could add an ID column, e.g.

```{r add-data-id}
 penguins %>%
  mutate(id = seq(1:nrow(penguins)))
```

This takes advantage of the `nrow()` function. Can you guess what it returns?

#### Exercise 4:

Compute the average bill length, bill depth, flipper length and body mass across all islands.

#### Exercise 5:

Is every species on every island?

---

## Summary statistics

What is a statistic? It's **any mathematical function of the data**. Sometimes, a statistic is referred to as "sample statistic" since you compute it from a finite **sample** (the data) and not the entire **population**.

For example, `penguins` is a sample of penguins in Antarctica, not an exhaustive list of the entire population.

Examples of statistics:

#### measure of central tendency:
- mean
- median
- mode

#### measures of spread:
- standard deviation
- variance
- range
- quartiles
- inter-quartile range (IQR)

#### order statistics:
- quantiles
- minimum (0 percentile)
- median (50th percentile)
- maximum (100 percentile)

... and any other arbitrary function of the data you can come up with!

#### Exercise 6: 

Come up with your own statistic and write it in the narrative below.

To access a column of the data, we'll use `data$column`.

Let's write down the R function for each of the above. 

#### Exercise 7:

Try to compute the above statistics for the penguin bill length column.

Do you receive an error? Why?

```{r compute-statistics}
# code here
```

Let's take a look at the **distribution** of bill length among all penguins.

```{r bl-plot-1}
penguins %>% # data
  ggplot(aes(x = bill_length_mm)) + # columns we want to look at
  geom_histogram() # geometry of the visualization
```

Let's visualize some of our summary statistics on the plot. 

#### Exercise 8: 
Uncomment the code below and fill in the blank with the mean.

```{r visualize-summaries, eval=F}
penguins %>% 
  ggplot(aes(x = bill_length_mm)) + 
  geom_histogram() #+ 
  #geom_vline(xintercept = __, color = 'red')
```

#### Exercise 9

Add another `geom_vline` with the median and mode. Use separate colors for each.


#### Exercise 10

Finally, let's try out another geometry `geom_density` and clean up our graph with some axes labels.

```{r density-plot}
# live coding here
```


