---
title: "Exploratory Data Analysis I"
subtitle: "STA 101"
format: html
---

## Bulletin

- [Lab 1](https://sta101-fa22.netlify.app/static/labs/lab01.html) due Thursday at 11:59pm on gradescope
- How to render to pdf directly

## Today

We'll begin today by visiting the last figure of [ae2](https://sta101-fa22.netlify.app/static/appex/ae2.html).

By the end of today you will...

- harness the power of `filter()` using logic
- create and interpret scatter plots, bar plots, stacked bar plots, facet plots and be able to look up and use other `ggplot` geometries


## Getting started

Download this application exercise by pasting the code below into your **console**

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae3.qmd",
destfile = "ae3.qmd")
```

## Load packages

```{r load-library, message = F, warning = F}
library(tidyverse)
library(palmerpenguins)
library(viridis) # we'll use to customize colors
```

## Load data

```{r load-data, message = F, warning = F}
data(penguins)
```

Type `?palmerpenguins` to learn more about this package. Or better yet, check it out [here](https://allisonhorst.github.io/palmerpenguins/).

## Logic in R

The table of logical operators below will be helpful as you work with filtering.

|operator | definition |
|:----------|:--------|
| `<` |   is less than? |
| `<=` |   is less than or equal to? |
| `>` |   is greater than? |
| `>=` |   is greater than or equal to? |
| `==` |   is exactly equal to? |
| `!=` |   is not equal to? |
| `is.na(x)` |   is x NA? |
| `!is.na(x)` |   is x not NA? |
| `x %in% y` |   is x in y? |
| `!(x %in% y)` |   is x not in y? |
| `x & y` |   is x AND y? |
| `x \| y` |   is x OR y? |
| `!x` |   is not x? |

The above operations return `TRUE` (1) or `FALSE` (0).

### Examples

How many penguins have flipper length > 200 mm?

```{r penguin-filter-1}
penguins %>%
  filter(flipper_length_mm > 200)
```

- We could also pipe into `nrow()` to quickly grab the number of rows. Try it!

How many **female** penguins have flipper length > 200 mm?

```{r penguin-filter-2}
penguins %>%
  filter(flipper_length_mm > 200 & (sex == "female"))
```

For how many penguins was flipper length not measured (i.e. reported as NA)?

```{r penguin-filter-3}
penguins %>%
  filter(is.na(flipper_length_mm))
```

How many penguins are of species Adelie *or* Chinstrap?

```{r}
penguins %>%
  filter(species %in% c("Adelie", "Chinstrap"))
```

#### Exercise 1:

Double check the *Adelie* and *Chinstrap* total using `count()`.

```{r}
# code here
```


#### Exercise 2:

What is the mean bill length of *Adelie* penguins? Hint: filter and then use `summarize()` as seen in lab 1.

```{r}
# code here
```


#### Exercise 3:

How many penguins have a bill length greater than 40 mm *or* a bill depth less than 15 mm?

```{r}
# code here
```

#### Exercise 4: 

What proportion of penguins are from the island `Torgersen`?

```{r}
# code here
```


## Plots

The procedure used to construct plots can be summarized using the code below.

```{r eval = FALSE}
ggplot(data = [data set], 
       mapping = aes(x = [x-variable], y = [y-variable])) +
   geom_xxx() +
   geom_xxx() + 
  other options
```

### Example: bar plot

```{r ex-bar-plot}
ggplot(data = penguins, 
       mapping = aes(x = species)) +
  geom_bar() +
  labs(x = "Species", y = "Count", title = "Palmer penguin species")
```

### Example: stacked bar plot

```{r bar-plot-fill}
penguins %>%
  filter(!is.na(sex)) %>%
ggplot(mapping = aes(x = species, fill = sex)) +
  geom_bar(position = "fill") +
  labs(x = "Species", y = "Count", title = "Palmer penguin species")
```

- try with and without `position = "fill"`

### Aesthetics

An aesthetic is a visual property in your plot that is derived from the data.

- shape
- color
- size
- alpha (transparency)

We can map a variable in our data set to a color, a size, a transparency, and so on. The aesthetics that can be used with each `geom_` can be found in the documentation.

Here we are going to use the viridis package, which has more color-blind accessible colors. `scale_color_viridis` specifies which colors you want to use. You can learn more about the options [here](https://www.rdocumentation.org/packages/viridis/versions/0.5.1/topics/scale_color_viridis).

Other sources that can be helpful in devising accessible color schemes include [Color Brewer](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3), the [Wes Anderson package](https://github.com/karthik/wesanderson), and the [cividis package](https://github.com/marcosci/cividis).

This visualization shows a scatterplot of bill length (x variable) and flipper length (y variable). Using the viridis function, we make points for male penguins purple and female penguins yellow. We also add axes labels and a title.

```{r aesthetic-1}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm,
                     color = sex)) + 
   geom_point() + 
   labs(title = "Flipper length vs bill length",
        x = "Bill length (mm)", y = "Flipper length (mm)") + 
        scale_color_viridis(discrete=TRUE, option = "D", name="Sex")
```

#### Exercise 5:
Can you remove the NAs from the above visualization?

**Question:** What will the visualization look like below? Write your answer 
down before running the code.

```{r aesthetic-2, eval = F}
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm,
                     shape = sex)) + 
   geom_point() + 
   labs(title = "Flipper length vs bill length",
        x = "Bill length (mm)", y = "Flipper length (mm)") + 
        scale_color_viridis(discrete=TRUE, option = "D", name="Sex")
```

### Faceting

We can use smaller plots to display different subsets of the data using faceting. This is helpful to examine conditional relationships.

```{r faceting-1}
penguins %>%
  ggplot(aes(x = bill_length_mm, flipper_length_mm, color = island)) +
  geom_point() +
  facet_wrap(~ species) +
  labs(x = "Bill length (mm)", y = "Flipper length (mm)", color = "Island")
```

```{r faceting-2}
penguins %>%
  ggplot(aes(x = bill_length_mm, flipper_length_mm, color = species)) +
  geom_point() +
  facet_wrap(~ island) +
  labs(x = "Bill length (mm)", y = "Flipper length (mm)", color = "Island") +
  scale_color_viridis(discrete = TRUE)
```

## `ggplot` activity

```{r ggplot-activity}
# code here
```

## Additional resources

- Find more ggplot geometries at [https://ggplot2.tidyverse.org/reference/](https://ggplot2.tidyverse.org/reference/)

