---
title: "Simple Regression"
subtitle: "STA 101"
format: html
---

## Bulletin

- lab 02 due tonight
- exam 01 next week

## Today

By the end of today you will...

- understand the grammar of linear modeling, including $y$, $x$, $\beta$, $\epsilon$, fitted estimates and residuals
- add linear regression plots to your 2D graphs
- be able to write a simple linear regression model mathematically and 
- fit the model to data in R in a `tidy` way

## Getting started

Download this application exercise by pasting the code below into your **console** (bottom left of screen)

```
download.file("https://sta101-fa22.netlify.app/static/appex/ae6.qmd",
destfile = "ae6.qmd")
```

## Load packages and data

```{r load-libraroies, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

Today's data is Apple and Microsoft stock prices from January 1st 2020 to December 31st 2021. I pulled this data off the Yahoo finance using their API via the `tidyquant` package July 2022.

```{r load-data, warning = F, message = F}
stocks = read_csv("https://sta101.github.io/static/appex/data/stocks1.csv")
```

## Notes

### The simple regression model and notation

$$
y = \beta_0 + \beta_1 x + \epsilon
$$

- $y$: the **outcome** variable. Also called the "response" or "dependent variable". In prediction problems, this is what we are interested in predicting.

- $x$: the **predictor**. Also commonly referred to as "regressor", "independent variable", "covariate", "feature", "the data".

- $\beta_0$, $\beta_1$ are called "constants" or **coefficients**. They are fixed numbers. These are **population parameters**. $\beta_0$ has another special name, "the intercept".

- $\epsilon$: the **error**. This quantity represents observational error, i.e. the difference between our observation and the true population-level expected value: $\beta_0 + \beta_1 x$.

Effectively this model says our data $y$ is linearly related to $x$ but is not perfectly observed due to some error.

### A simple example

Let's examine January 2020 open prices of Microsoft and Apple stocks to illustrate some ideas.

```{r simple-example}
stocks_subset = stocks %>%
  slice(1:21)

stocks_subset %>%
  ggplot(aes(x = MSFT.Open, y = AAPL.Open)) +
  geom_point() + 
  labs(x = "MSFT Open", y = "AAPL Open", title = "Open prices of MSFT and AAPL January 2020") +
  theme_bw() 
  # more code here
```

#### Exercise 1

Add `geom_abline()` to the above plot and try different slopes and intercepts until you find a trendline you are satisfied with. The equation below describes your **fitted model**. Re-write the equation below, filling in $\hat{\beta_0}$ and $\hat{\beta_1}$ with your estimates.

$$
\hat{y} = \hat{\beta_0} +  \hat{\beta_1} x
$$

- $\hat{y}$ is the **expected outcome**.
- $\hat{\beta}$ is the **estimated** or **fitted** coefficient
- there is no error term here because we do not predict error

The equation of my line above:

$$
\text{[your equation here]}
$$

The central idea is that if we measure every $x$ and every $y$ in existence, ("the entire population") there is some true "best" $\beta_0$ and $\beta_1$ that describe the relationship between $x$ and $y$. Since we only have a **sample** of the data, we estimate $\beta_0$ and $\beta_1$. We call our estimates $\hat{\beta_0}$, $\hat{\beta_1}$ "beta hat". We never have all the data, thus we never can really know what the true $\beta$s are.

## Ordinary least squares (OLS) regression

### The residuals

For any linear equation we write down, there will be some difference between the predicted outcome of our linear model ($\hat{y}$) and what we observe ($y$)... (But of course! Otherwise everything would fall on a perfect straight line!)

This difference between what we observe and what we predict $y - \hat{y}$ is known as a residual $r$.

More concisely,

$$
r = y - \hat{y}
$$

Residuals are dependent on the line we draw. Visually, here is a model of the data, $y = -5 + \frac{1}{2}x$ and 1 of the residuals is outlined in red.

```{r, echo = F}

# day11 = c(164.35, 78.4) # data point

stocks_subset %>%
  ggplot(aes(x = MSFT.Open, y = AAPL.Open)) +
  geom_point() + 
  labs(x = "MSFT Open", y = "AAPL Open", title = "Open prices of MSFT and AAPL January 2020") +
  theme_bw() +
  geom_abline(slope = 0.5, intercept = -5) +
  geom_segment(x = 164.35, xend = 164.35, y  = 78.4, yend = 77.175, color = 'red')
```

There is, in fact, a residual associated with every single point in the plot.

```{r all-residuals}
predictAAPL = function(x) {
  return(-5 + (0.5*x))
}

xPoints = stocks$MSFT.Open[1:21]
yPoints = stocks$AAPL.Open[1:21]
yHat = predictAAPL(xPoints)

stocks_subset %>%
  ggplot(aes(x = MSFT.Open, y = AAPL.Open)) +
  geom_point() + 
  labs(x = "MSFT Open", y = "AAPL Open", title = "Open prices of MSFT and AAPL January 2020") +
  theme_bw() +
  geom_abline(slope = 0.5, intercept = -5) +
  geom_segment(x = xPoints, xend = xPoints, y  = yPoints, yend = yHat, color = 'red')


```

We often wish to find a line that fits the data "really well", but what does this mean? Well, we want small residuals! So we pick an **objective function**. That is, a function we wish to minimize or maximize.

### The objective function

#### Exercise 2

At first, you might be tempted to minimize $\sum_i r_i$, but this is problematic. Why? Can you come up with a better solution (other than the one listed below)?

[answer here]

In practice, we minimize the **sum of squared residuals**: 

$$
\sum_i r_i^2
$$

Note, this is the same as 

$$
\sum_i (y_i - \hat{y})^2
$$

#### Exercise 3

Check out an interactive visualization of "least squares regression" [here](https://seeing-theory.brown.edu/regression-analysis/index.html#section1). Click on `I` and drag the points around to get started. Describe what you see.

[response here]

#### Exercise 4

- Check for understanding

How far off is your model (from exercise 1) from the actual observed data on January 11 2020? The observed value is  MSFT: \$164.35 and AAPL: \$78.4. Compute the single square residual using your **fitted model** from exercise 1.

```{r compute-a-residual}
# code here
```

## Plotting the OLS regression line

Plotting the OLS regression line, that is, the line that minimizes the sum of square residuals is very easy with ggplot. Simply add 

```
geom_smooth(method = 'lm', se = F)
```

to your plot.

`method = lm` says to draw a line according to a "linear model" and `se = F` turns off standard error bars. You can try without these options for comparison.

Optionally, you can change the color of the line, e.g.

```
geom_smooth(method = 'lm', se = F, color = 'red')
```


#### Exercise 5 

Copy your code from exercise 1 below. Add `geom_smooth()` as described above with `color = 'steelblue'` to see how close your line is.

```{r how-close-ex-5}
# code here
```

## Finding $\hat{\beta}$

To **fit the model** in R, i.e. to "find $\hat{\beta}$", use the code below as a template:

```
modelFit = linear_reg() %>%
  set_engine("lm") %>%
  fit(y-variable-here ~ x-variable-here, data = data-frame-here)
```

- `linear_reg` tells `R` we will perform linear regression
- `set_engine` tells `R` to use the standard linear modeling (lm) machinery
- `fit` defines the outcome $y$, predictor $x$ and the data set

Running the code above, but replacing the arguments of the `fit` command appropriately will create a new object called "modelFit" (defined on the first line) that stores all information about your fitted model.

To access the information, you can run, e.g.

```
tidy(modelFit)
```

Let's try it out.

#### Exercise 6

Find the OLS fitted linear model $y = \hat{\beta_0} + \hat{\beta_1} x$ for January 2020, where $x$ is Microsoft's opening price and $y$ is Apple's opening price. Print your results to the screen

```{r fit-MSFT-AAPL-January}
# code here
```


#### Exercise 7

Re-write the fitted equation replacing $\beta_0$ and $\beta_1$ with the OLS fitted values.

$$
\text{[your equation here]}
$$