---
title: "Untitled"
output: html_document
date: '2022-06-30'
---

```{r}
library(tidyverse)
library(knitr)
```

## Notes

```{r likelihood-1}
p = seq(0,1,0.001)
likelihood = dbinom(53, 78, p)
df = data.frame(p, likelihood)

df %>%
  ggplot(aes(x = p, y = likelihood)) +
  geom_point() + 
  geom_vline(xintercept = 0.6794872)
```

```{r OLS-objective}
courage %>%
  mutate(home = ifelse(home_team == "NC", "home", "away"),
         result2 = ifelse(result == "win", "win", "not_win")) %>%
   count(result2, home) %>%
  kable()

away_win = 22
home_win = 31
away_not_win = 14
home_not_win = 11

home_total = home_win + home_not_win
away_total = away_win + away_not_win

likelihood2 = function(p, q) {
  return(
    dnorm(home_win, home_total, p) * dnorm(away_win, away_total, q)
  )
}

p = seq(0,1,0.1)
q = p
p = rep(p, each = length(p))
q = rep(q, length(q))

likelihood = likelihood2(p,q)

df = data.frame(p, q, likelihood)

df %>%
ggplot(aes(x = p, y = q, fill = likelihood)) +
  geom_tile()
```



Today's data comes from the Flint, Michigan water crisis. See [The murky tale of Flint's deceptive water data](https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2017.01016.x) for further reading. A quote from the article: "This is a cautionary tale of how flawed and inadequate testing creates misleading data".

```{r load-data, warning = F, message = F}
flint = read_csv("~/Desktop/teaching/sta101/static/static/appex/data/flint.csv")
```

### Codebook

- **id**: sample ID number
- **zip**: ZIP code in Flint of the sample's location
- **ward**: ward in Flint of the sample's location
- **draw**: which time point the water was sampled from
- **lead**: lead content in parts per billion

## Notes


### Poor water testing

```{r first-look-at-data, echo = F}
flint %>%
  filter(draw == "third") %>%
  group_by(zip) %>%
  summarize(mean(lead))
```


```{r}
flint %>%
  ggplot(aes(x = id, y = lead, color = draw)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylim(c(0,50))
```


### Main effects model

$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \epsilon
$$

### Interactive effects model

$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1 x_2 + \epsilon
$$

```{r flint-1}
flint3 %>%
  filter(Lead_ppb <= 10) %>%
  filter(SampleLocation %in% c("BATHROOM", "KITCHEN")) %>%
  ggplot(aes(x = SampleNumber, y = Lead_ppb, color = SampleLocation)) +
  geom_point()+
  geom_smooth(method = 'lm')
```
