---
title: "STA101 Schedule"
subtitle: Fall, 2022
output:
  html_document:
    css: custom-site.css
    df_print: paged
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval=F)
```

```{r load-libraries}
library(tidyverse)
library(knitr)
library(emo)
```

```{r smart-date-setup}
# set these:
firstDay = as.Date("2022-08-29")
firstDayID = "Mon"
lastDay = as.Date("2022-12-09")
lectureDays = c("Tue", "Thu")
labDays = c("Fri")
holidays = as.Date(c("2022-10-10", "2022-10-11", "2022-11-23", "2022-11-24", "2022-11-24"))
### render table
allDates = seq(from = firstDay, to = lastDay, by = 'day')
schedule = data.frame(allDates,
                        dayOfWeek = weekdays(allDates, abbreviate=T)) %>%
  filter(dayOfWeek %in% c(lectureDays, labDays))
schedule = schedule %>%
  mutate(lab = ifelse(dayOfWeek %in% labDays, TRUE, FALSE)) %>%
  mutate(Topic = ifelse(allDates %in% holidays, "NO CLASS", "")) %>%
  mutate(Date = format(allDates, format = "%b %d")) %>%
  mutate(Date = paste(dayOfWeek, Date)) %>%
  mutate(Week = rep("", nrow(schedule))) %>%
  mutate(Prepare = rep("", nrow(schedule))) %>%
  mutate(AE = rep("", nrow(schedule))) %>%
  mutate(Assignment = rep("", nrow(schedule))) %>%
  mutate(Exam = rep("", nrow(schedule))) %>%
  mutate(Project = rep("", nrow(schedule))) %>%
  select(Week, Date, Topic, Prepare, AE, Assignment, Exam, Project)
# to be pasted into console then pasted into markdown:
schedule %>%
  kable(format = "markdown")
```

Find all assignments and due dates here. This schedule updates frequently. 

|Week |Date       |Topic    |Prepare |AE |Lab |Exam |Project |
|:----|:----------|:--------|:-------|:----|:--------|:----|:-------|
|  1   |Tue Aug 30 | Welcome to STA 101!        |  [`r emo::ji("book")`](/prepare/week-1)      | [ae1](/static/appex/ae1.html), \ \ [`r emo::ji("computer")`](/static/slides/day01.html)   |           |     |        |
|     |Thu Sep 01 | Intro to R       | [`r emo::ji("book")`](/prepare/week-1)        |   |         |     |        |
|     |Fri Sep 02 | Hello R!      |        |   |  lab 1         |     |        |
|  2   |Tue Sep 06 | Exploratory data analysis I         |        |   |           |     |        |
|     |Thu Sep 08 |  Exploratory data analysis II       |        |   |           |     |        |
|     |Fri Sep 09 |   EDA      |        |   | lab 2          |     |        |
| 3    |Tue Sep 13 | Exploratory data analysis III         |        |   |           |     |        |
|     |Thu Sep 15 | Simple regression        |        |   |           |     |        |
|     |Fri Sep 16 | EDA and simple regression   |        |   |  lab 3         |     |        |
|  4   |Tue Sep 20 |  Multiple regression I       |        |   |           |     |        |
|     |Thu Sep 22 |  Multiple regression II       |        |   |           | exam 1    |        |
|     |Fri Sep 23 |   No lab: exam      |        |   |           |     |        |
|  5   |Mon Sep 26 |        |        |   |           | exam 1 due    |        |
|     |Tue Sep 27 |  Multiple regression III, Prediction       |        |   |           |     |  Regression project      |
|     |Thu Sep 29 |  Logistic regression       |        |   |           |     |        |
|     |Fri Sep 30 | All about regression   |        |   |  lab 4         |     |        |
|  6   |Tue Oct 04 |  Random variables      |        |   |           |     |        |
|     |Thu Oct 06 |   Probability    |        |   |           |     |        |
|     |Fri Oct 07 |  Project workday    |        |   |         |     |        |
|  7   |Tue Oct 11 |NO CLASS: Fall break |        |   |           |     |        |
|     |Thu Oct 13 |  Distributions I     |        |   |           |     |  Final project      |
|     |Fri Oct 14 |   Probability and distributions    |        |   |  lab 5         |     |        |
|  8   |Tue Oct 18 |  Distributions II    |        |   |           |     |        |
|     |Thu Oct 20 |   Sampling and the bootstrap    |        |   |           |     |        |
|     |Fri Oct 21 |    The bootstrap     |        |  |   lab 6         |     |        |
|  9   |Tue Oct 25 |  Central limit theorem I     |        |   |           |     |        |
|     |Thu Oct 27 |  Central limit theorem II    |        |   |           |     |        |
|     |Fri Oct 28 |   CLT and CI      |        |   |  lab 7         |     |        |
|  10   |Tue Nov 01 |  Hypothesis testing I       |        |   |           |     |        |
|     |Thu Nov 03 |  Hypothesis testing II      |        |   |           |     |        |
|     |Fri Nov 04 |    Hypothesis tests     |        |   |   lab 8         |     |        |
|  11   |Tue Nov 08 | Hypothesis testing III       |        |   |           |     |        |
|     |Thu Nov 10 | Hypothesis testing recap      |        |   |           | exam 2    |        |
|     |Fri Nov 11 |   No lab: exam released      |        |   |           |     |        |
|  12   |Mon Nov 14 |         |        |   |           | exam 2 due    |        |
|       |Tue Nov 15 | Data science ethics        |        |   |           |     |        |
|     |Thu Nov 17 |   Project report tips      |        |   |           |     |        |
|     |Fri Nov 18 | New packages     |        |  |   lab 9        |     |        |
|  13   |Tue Nov 22 | Project presentation tips        |        |   |           |     |        |
|     |Thu Nov 24 |NO CLASS: Thanksgiving |        |   |           |     |        |
|     |Fri Nov 25 |  NO CLASS: Thanksgiving       |        |   |           |     |        |
|  13   |Tue Nov 29 | Special topic: genetic forensic analysis        |        |   |           |     |        |
|     |Thu Dec 01 |  Special topic: cryptanalysis      |        |   |           |     |        |
|     |Fri Dec 02 |   Final project peer review     |        |   |           |     |        |
|  15   |Tue Dec 06 | Special topic        |        |   |           |     |        |
|     |Thu Dec 08 | Special topic      |        |   |           |     |        |
|     |Fri Dec 09 |  Project workday      |        |   |           |     |        |