---
title: "Lab 06"
author: "STA 101"
subtitle: "Due Thursday October 27 at 11:59pm"
format: html
css: "labs2.css"
---

By the end of this lab you will



## Getting started

\ 1. Download the lab template by pasting the code below in your **console**:

```
download.file("https://sta101-fa22.netlify.app/static/labs/lab06_template.qmd",
              destfile = "lab06.qmd")
```

\ 2. Under the "Files" tab on the right hand side, click on `lab06.qmd` to open the lab template.


\ 3. Complete the exercises below using the space provided.

## Warm up

Be sure to update the YAML at the top of the document to include your name and the date.

### Packages

```{r load-packages, warning=F, message=F}
library(tidyverse)
library(openintro)
library(tidymodels)
```

### Data

Load the data:

```{r load-data, warning=F, message=F}
data(bob_ross)
```

Bob Ross was a painter who was most famous for his PBS television show The Joy of Painting. In each episode, Ross created a new oil painting and provided instructions and commentary as he painted it. Ambitious viewers could paint along but viewers also simply enjoyed watching and listening to Ross's soothing voice as he painted an outdoor scene in 30 minutes.

In 2014, Walt Hickey wrote an article for FiveThirtyEight using statistics to analyze the paintings created on the show. The article focused on features that were often seen in Ross’s paintings, such as trees, clouds, cabins, among others. [Click here to see the article](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/)

In this assignment, you will analyze the data that was used for the article. The data is in the `bob_ross` data set in the fivethirtyeight R package. Each observation represents an episode of the TV show. One painting was created in an episode. To access the full codebook of variables, explore the documentation using `?bob_ross`.

We'll focus on the following variables in this assignment:

- `cabin`: Whether the painting contains a cabin
- `cirrus`: Whether the painting contains cirrus clouds
- `conifer`: Whether or not a conifer tree is present
- `cumulus`: Whether the painting contains cumulus clouds
- `guest`: Whether or not the episode featured a guest painter
- `lake`: Whether the painting contains a lake
- `season`: Season the painting is from
- `snowy_mountain`: Whether or not the a snowy mountain is present
- `steve_ross`: Whether or not Steve Ross was the featured guest
- `tree`: Whether or not the painting contains at least 1 tree

:::{.callout-important}
Render as you work! Render at least before and after each exercise to avoid last minute rendering issues.
:::

## Exercises

*"There’s nothing wrong with having a tree as a friend."*

1. How many episodes are there of the show *The Joy of Painting*? In how many episodes was at least one tree painted? What is the probability that a randomly selected episode features at least one tree? Given that an episode feature a tree, what is the probability the episode features a `conifer`?

*"Let's build us a happy, little cloud that floats around the sky."*

2. Are events **A** and **B** disjoint? Why or why not?

- **A**: A guest paints a `cirrus` cloud
- **B**: A guest paints a `cumulus` cloud

3. In the FiveThirtyEight article, Walt Hickey calculates various probabilities to describe the combination of features typically found in Bob Ross paintings. He states the following about the presence of cabins and lakes in Ross's paintings: "About 18 percent of his paintings feature a cabin. Given that **Bob Ross painted** a cabin, there's a 35 percent chance that it's on a lake..."

- How many of **Bob Ross's paintings** feature a cabin? Call this number "$k$"

- How many of those that feature a cabin also feature a lake? Call this number "$x$".

- Imagine the following: every time Bob Ross paints a cabin, he flips a **fair coin** to decide whether or not to paint a lake. Given a collection of $k$ Bob Ross paintings with cabins, what is the probability Bob Ross painted $x$ or fewer lakes? Hint: simulate this scenario using the code below as a template.

```{r simulation-code, eval = F}
set.seed(10142022) # don't change the seed
num_lakes = rbinom(10000, k, prob = ?)
cabin_lakes = data.frame(num_lakes)
```

4. Are events **C** and **D** independent? Why or why not?

- **C**: A painting features a snowy mountain
- **D**: A painting is from season 1

5. The **Joy of Painting** occasionally featured a guest painter other than Bob Ross. One guest painter was Bob's son Steve Ross. Imagine that before each episode, Bob Ross flips a coin to decide whether or not to invite his son Steve to be a guest painter. If the coin lands heads, Bob invites Steve. If Steve is available, he agrees to paint for an episode, but if he's busy, then Bob finds another guest painter.

Using the data set and the model described above, report your best guess of the following quantities:

- p(Steve is busy | Bob invites Steve)
- p(the coin Bob flips lands "Heads")

*"However you think it should be, that’s exactly how it should be."*

6. Your turn! Use this data to explore a question of your choice about paintings created in the TV show *The Joy of Painting*. Your question should explore the relationship between 3 variables in the data set; at least one of the variables must be one that hasn't been used in exercises 1 - 5. You may use the entire data set or focus the analysis on paintings made by Bob Ross.

- Clearly state the question you’re exploring.
- Create 1 - 2 visualizations that can be used to explore the question. Customize the colors of your visualization using some of the colors commonly used in Bob Ross paintings. [Click here](https://sites.radford.edu/~cdc/seansevillair.html) for a list of common Bob Ross hex codes.

:::{.callout-hint}
[Click here for functions to manually create color palettes in ggplot2.](https://ggplot2.tidyverse.org/reference/scale_manual.html)
:::
    
- Calculate the relevant probabilities needed to answer your question.
- Write a short narrative, about 2 - 4 sentences, answering the research question based on the visualization and probabilities. Note: The narrative should not exceed 4 sentences.

*"Anything that you're willing to practice, you can do."*

7. What does it mean that the "Normal distribution is **stable**"? Let $X \sim N(0, 3)$ and $Y \sim N(1, 2)$ where $X \sim N(\mu, \sigma)$ means "X is normally distributed with mean $\mu$ and standard deviation $\sigma$". Write down the distributions of the following (including location and scale parameters):

- $X + Y$
- $\frac{1}{2}X - 2Y$
- $X - 2X + Y$

Note: no code is required for exercise 7.

8. Plot a histogram of the `num_lakes` from exercise 3. Overlay the plot with a normal distribution with  appropriate mean $\mu$ and standard deviation $\sigma$ to approximate the histogram. If $X \sim N(\mu, \sigma)$, what is $p(X \leq x)$? Recall that we define lower-case $x$ in exercise 3. How does this probability compare to your result from exercise 3?


## Formatting

**Reminder**: For all assignments in this course there is a "formatting" component to the grade. To receive full points for "formatting", you must:

\ 1. Have your name at the top of the rendered document

\ 2. Pipes `%>%` and ggplot layers `+` should be followed by a newline (see formatting above)

\ 3. Your code should be under the 80 character code limit. (You shouldn't have to scroll to see all your code on the rendered document).

\ 4. All exercises and corresponding pages should be linked on gradescope.

These necessary "tidyverse" style choices are good general practice and will help make your code more legible. For a more extensive list of recommended guidelines, [click here](https://style.tidyverse.org/).

## Submitting to gradescope

Once you are fully satisfied with your lab, render to .pdf to create a .pdf document. You may notice that the formatting/theme of the report has changed – this is expected. Remember – you must turn in a .pdf file to the Gradescope page before the submission deadline for credit. To submit your assignment: 

- Go to [http://www.gradescope.com](http://www.gradescope.com) and click Log in in the top right corner. - Click `School Credentials`, `Duke NetID` and log in using your NetID credentials. 

- Click on your STA 101 course. 

- Click on the assignment, and you'll be prompted to submit it. 

- Mark the pages associated with each exercise. All of the papers of your lab should be associated with at least one question (i.e., should be “checked”). Select the first page of your .pdf submission to be associated with the "Formatting" section.

## Grading

Total: 50 pts.

    Exercise 1: 5pts

    Exercise 2: 3pts

    Exercise 3: 5pts

    Exercise 4: 7pts

    Exercise 5: 4pts
    
    Exercise 6: 8pts
    
    Exercise 7: 6pts
    
    Exercise 8: 8pts
    
    Workflow and formatting:  4pts
    