---
title: "Schedule"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval=F)
```

```{r load-libraries}
library(tidyverse)
library(knitr)
```

```{r smart-date-setup}
# set these:
firstDay = as.Date("2022-01-07")
firstDayID = "Fri"
lastDay = as.Date("2022-04-20")

lectureDays = c("Wed", "Fri")
labDays = c("Tue")
holidays = as.Date(c("2022-03-08", "2022-03-09", "2022-03-11"))

### render table

allDates = seq(from = firstDay, to = lastDay, by = 'day')
schedule = data.frame(allDates,
                        dayOfWeek = weekdays(allDates, abbreviate=T)) %>%
  filter(dayOfWeek %in% c(lectureDays, labDays)) %>%
  mutate(lab = ifelse(dayOfWeek %in% labDays, TRUE, FALSE)) %>%
  mutate(Topic = ifelse(allDates %in% holidays, "NO CLASS", "")) %>%
  mutate(Date = format(allDates, format = "%b %d")) %>%
  mutate(Date = paste(dayOfWeek, Date)) %>%
  mutate(Week = rep("", nrow(schedule))) %>%
  mutate(Prepare = rep("", nrow(schedule))) %>%
  mutate(AE = rep("", nrow(schedule))) %>%
  mutate(Assignment = rep("", nrow(schedule))) %>%
  mutate(Exam = rep("", nrow(schedule))) %>%
  mutate(Project = rep("", nrow(schedule))) %>%
  select(Week, Date, Topic, Prepare, AE, Assignment, Exam, Project)


# to be pasted into console then pasted into markdown:
schedule %>%
  kable(format = "markdown")
```

|Week |Date       |Topic    |Prepare |AE |Assignment |Exam |Project |
|:----|:----------|:--------|:-------|:--|:----------|:----|:-------|
|     |Fri Jan 07 |         |        | [ae1](/static/static/appex/ae1.html)  |           |     |        |
|     |Tue Jan 11 |         |        |   |           |     |        |
|     |Wed Jan 12 |         |        |   |           |     |        |
|     |Fri Jan 14 |         |        |   |           |     |        |
|     |Tue Jan 18 |         |        |   |           |     |        |
|     |Wed Jan 19 |         |        |   |           |     |        |
|     |Fri Jan 21 |         |        |   |           |     |        |
|     |Tue Jan 25 |         |        |   |           |     |        |
|     |Wed Jan 26 |         |        |   |           |     |        |
|     |Fri Jan 28 |         |        |   |           |     |        |
|     |Tue Feb 01 |         |        |   |           |     |        |
|     |Wed Feb 02 |         |        |   |           |     |        |
|     |Fri Feb 04 |         |        |   |           |     |        |
|     |Tue Feb 08 |         |        |   |           |     |        |
|     |Wed Feb 09 |         |        |   |           |     |        |
|     |Fri Feb 11 |         |        |   |           |     |        |
|     |Tue Feb 15 |         |        |   |           |     |        |
|     |Wed Feb 16 |         |        |   |           |     |        |
|     |Fri Feb 18 |         |        |   |           |     |        |
|     |Tue Feb 22 |         |        |   |           |     |        |
|     |Wed Feb 23 |         |        |   |           |     |        |
|     |Fri Feb 25 |         |        |   |           |     |        |
|     |Tue Mar 01 |         |        |   |           |     |        |
|     |Wed Mar 02 |         |        |   |           |     |        |
|     |Fri Mar 04 |         |        |   |           |     |        |
|     |Tue Mar 08 |NO CLASS |        |   |           |     |        |
|     |Wed Mar 09 |NO CLASS |        |   |           |     |        |
|     |Fri Mar 11 |NO CLASS |        |   |           |     |        |
|     |Tue Mar 15 |         |        |   |           |     |        |
|     |Wed Mar 16 |         |        |   |           |     |        |
|     |Fri Mar 18 |         |        |   |           |     |        |
|     |Tue Mar 22 |         |        |   |           |     |        |
|     |Wed Mar 23 |         |        |   |           |     |        |
|     |Fri Mar 25 |         |        |   |           |     |        |
|     |Tue Mar 29 |         |        |   |           |     |        |
|     |Wed Mar 30 |         |        |   |           |     |        |
|     |Fri Apr 01 |         |        |   |           |     |        |
|     |Tue Apr 05 |         |        |   |           |     |        |
|     |Wed Apr 06 |         |        |   |           |     |        |
|     |Fri Apr 08 |         |        |   |           |     |        |
|     |Tue Apr 12 |         |        |   |           |     |        |
|     |Wed Apr 13 |         |        |   |           |     |        |
|     |Fri Apr 15 |         |        |   |           |     |        |
|     |Tue Apr 19 |         |        |   |           |     |        |
|     |Wed Apr 20 |         |        |   |           |     |        |
